There are no small Advantages: Reinforcement learning in near continuous time with Deep RelAtive Mixture of Advantages

### 2019-01-14 paper/icml_drau/advup.tex
lots in framework, remarks in the other places

\subsection{Advantage Updating}
\label{subsec:reparam}
To define an object which contains the same information on actions as
$Q^\pi_\deltat$, but admits an action dependant continuous time limit, , it is
natural to define
\begin{align}
	A^\pi_\deltat(s, a) &= \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat},
    \label{eq:adv}
\end{align}
a rescaled version of the advantage function, as the difference between between
$Q^\pi_\deltat(s, a)$ and $V^\pi\deltat(s)$ is of order $\bigO(\deltat)$.
Contrary to the action state value function, the rescaled advantage function converges
to an action dependent quantity when $\deltat$ goes to $0$
\begin{align}
	A^\pi_\deltat(s, a)=&\reward(s, a)+\frac{\gamma^{\deltat}\E_{\delta s \sim \delta s(s, a)}V^\pi_\deltat(s + \delta s)-V^\pi\deltat(s)}{\deltat}\nonumber\\
	=&\reward(s, a)+\log(\gamma) V^\pi(s)+\partial_s V^\pi(s)f(s, a)\nonumber\\
         &+\frac{\text{tr}\left(\Sigma^T(s, a)\partial^2_{s^2}V^\pi(s) \Sigma(s,a)\right)}{2} + O(\deltat)\nonumber.
\end{align}
The discretized state action value function can then be rewritten as
\begin{equation}
	Q^\pi_\deltat(s, a) = V^\pi_\deltat(s) + \deltat A^\pi_\deltat(s, a).
	\label{eq:reparam_q_pi}
\end{equation}
A natural way of approximating $V^\pi_\delta$ and $A^\pi_\deltat$ is to apply
\emph{Sarsa} or Q-learning to a reparameterized Q-function approximator
\begin{equation}
	Q_\Theta(s, a) = V_{\theta}(s) + \deltat A_{\psi}(s, a).
\end{equation}
To avoid cumbersome notations, in what follows, parameter indices are dropped.
At initialization, if both $V_{\theta}$ and $A_{\psi}$ are initialized
independantly of $\deltat$, this parameterization reasonably scales the
contribution of actions relative to states in the value of $Q$.

@@

A sufficient condition for the exploratory policy to admit a time continuous
limit is for $\pi^\text{explore}_\deltat$ to be a measurable function, and for
$\nu_\deltat$ to admit one, i.e. that the sequence $\nu_\deltat$ converges to a
well defined continuous stochastic process $\nu$ as $\deltat$ goes to $0$.
\TODO{I'm convinced that this is true, reviewers might not be.}

The exploration scheme used~\cite{ddpg}, which consists in adding an
\emph{Ornstein-Uhlenbeck}~\cite{orn-uhl} process (OU process) to actions, is naturally time.
invariant. It can be rewritten as
\begin{equation}
	a^k_\deltat = \pi(s^k_\deltat) + \nu^k_\deltat
\end{equation}
where $\nu^k_\deltat$ is the discretization of a time continuous \emph{Ornstein-Uhlenbeck}
process, defined as
\begin{equation}
	d\nu_t = - \nu_t \theta dt + \sigma \dbt.
	\label{eq:orn_uhl}
\end{equation}

@@

\subsection{Algorithms}
\label{subsec:algorithm}
To learn $V_{\theta}$ and $A_{\psi}$, variants of Q-learning for continuous~\cite{ddpg} and
discrete~\cite{dqn} action spaces are used. In details, $V_{\theta}$ and $A_{\psi}$ are learnt
using approximate version of the following bellman equation, which holds for the true $A^\pi_\deltat$
and $V^\pi_\deltat$,
\begin{align}
	V^{\pi}_\deltat(s) + \deltat A^{\pi}_\deltat(s, a) &= r\deltat + \gamma^{\deltat}  \E_{s'} V^{\pi}_\deltat(s')\label{eq:bellman_A}\\
	A^{\pi}_\deltat(s, \pi(s)) &= 0\label{eq:max_A}
\end{align}
for a greedy exploitation policy $\pi(s) = \text{argmax}_{a'}A^\pi_\deltat(s,
a')$, exactly implemented in the case of discrete actions, and approximated by
a neural network $\pi_\phi(s)$, trained to maximize (an approximation of)
$A^\pi_\deltat(s, \pi_\phi(s))$ in the continuous case.  Eq.~\eqref{eq:max_A}
is directly verified by $A_{\psi}$, owing to the reparameterization
$A_\psi(s, a) = \bar{A}_\psi(s, a) - \bar{A}_\psi(s, \pi(s))$, described
in~\ref{subsec:reparam}.  To approximately verify~\eqref{eq:bellman_A}, the
corresponding squared residual is minimized by an approximate gradient descent.
The corresponding update equations when learning from a transition $(s, a, r, s')$ of
an exploratory trajectory are
\begin{align}
	\delta Q_\deltat &\leftarrow A(s, a) \deltat - \left(r \deltat + \gamma^{\deltat} V(s') - V(s)\right)\\
	\theta_\deltat &\leftarrow \theta_\deltat + \eta^1_\deltat \partial_{\theta} V(s) \frac{\delta Q_\deltat}{\deltat}\\
	\psi_\deltat &\leftarrow \psi_\deltat + \eta^2_\deltat \partial_{\psi} A(s, a) \frac{\delta Q_\deltat}{\deltat}.
	\label{eq:approx_bellman_A}
\end{align}
where $\eta$'s are learning rates. As we will see in the following section,
to obtain a well defined continuous time limit, the learning rates can be rewritten as
$\eta^i_\deltat = \alpha^i \deltat$, where $\alpha^1$ and $\alpha^2$ are independant
of $\deltat$. In practice, $A$ and $V$ are learned using
transitions sampled from a replay buffer instead of trajectory samples, to
prevent excessive correlation between training samples~\cite{dqn}.
\begin{algorithm}[ht]
	\caption{Deep Advantage Updating (Discrete actions)}
	\input{dauc.tex}
	\label{alg:dau}
\end{algorithm}

>>>>

\subsection{Advantage Updating}
\label{subsec:reparam}
To define an object which contains the same information on actions as
$Q^\pi_\deltat$, but admits an action dependant continuous time limit, it is
natural to define
\begin{align}
	A^\pi_\deltat(s, a) &= \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat},
    \label{eq:adv}
\end{align}
a rescaled version of the advantage function, as the difference between between
$Q^\pi_\deltat(s, a)$ and $V^\pi\deltat(s)$ is of order $\bigO(\deltat)$.
Contrary to the action state value function, the rescaled advantage function converges
to an action dependent quantity when $\deltat$ goes to $0$ \LB{In next equation, tiny detail but I would write $\frac{1}{\deltat}(r\deltat+...)$, easier to read.}
\begin{align}
	A^\pi_\deltat(s, a)=&\reward(s, a)+\frac{\gamma^{\deltat}\E_{\delta s \sim \delta s(s, a)}V^\pi_\deltat(s + \delta s)-V^\pi_\deltat(s)}{\deltat}\nonumber\\
	=&\reward(s, a)+\log(\gamma) V^\pi(s)+\partial_s V^\pi(s)f(s, a)\nonumber\\
         &+\frac{\text{tr}\left(\Sigma^T(s, a)\partial^2_{s^2}V^\pi(s) \Sigma(s,a)\right)}{2} + O(\deltat)\nonumber.
\end{align}

\LB{It might be my math bias, but I would really prefer to read something like:
  Contrary to the action state value function, the rescaled advantage function converges
  to an action dependent quantity when $\deltat$ goes to $0$, which contains all the necessary information to improve the strategy. We have the following theorem:
  \begin{theorem}
    There is function $A^\pi(s, a)$ such that for all $a, s$, we have:
    \begin{equation}
      A^\pi_\deltat(s, a) \rightarrow_{\deltat \rightarrow 0} A^\pi(s, a)
    \end{equation}
    Moreover, $A^\pi$ contains all the necessary information about actions. We have $\pi$ is the optimal strategy if and only if $A^\pi(s, \pi(s)) = 0$ for all $s \in {\cal S}$.
    [[I don't know if this is true, but this is the kind of property we want for A, and this is the idea of not only it converges, but it converges to a meaningful limit]]
  \end{theorem}
  And the equation which is there before would only be in the appendix, in the proof.
}
The discretized state action value function can then be rewritten as
\begin{equation}
	Q^\pi_\deltat(s, a) = V^\pi_\deltat(s) + \deltat A^\pi_\deltat(s, a).
	\label{eq:reparam_q_pi}
\end{equation}
A natural way of approximating $V^\pi_\delta$ and $A^\pi_\deltat$ is to apply
\emph{Sarsa} or Q-learning to a reparameterized Q-function approximator
\begin{equation}
	Q_\Theta(s, a) = V_{\theta}(s) + \deltat A_{\psi}(s, a).
\end{equation}
To avoid cumbersome notations, in what follows, parameter indices are dropped.
At initialization, if both $V_{\theta}$ and $A_{\psi}$ are initialized
independantly of $\deltat$, this parameterization reasonably scales the
contribution of actions relative to states in the value of $Q$.

@@

\LB{All this subsection is not extremely clear to me, especially the next paragraph. And it is a detail, but I'm not sure I like $\nu$ as a notation for a random variable.} A sufficient condition for the exploratory policy to admit a time continuous
limit is for $\pi^\text{explore}_\deltat$ to be a measurable function, and for
$\nu_\deltat$ to admit one, i.e. that the sequence $\nu_\deltat$ converges to a
well defined continuous stochastic process $\nu$ as $\deltat$ goes to $0$.
\TODO{I'm convinced that this is true, reviewers might not be.}

The exploration scheme used~\cite{ddpg}, which consists in adding an
\emph{Ornstein-Uhlenbeck}~\cite{orn-uhl} process (OU process) to actions, is naturally time
invariant. It can be rewritten as
\begin{equation}
	a^k_\deltat = \pi(s^k_\deltat) + \nu^k_\deltat
\end{equation}
where $\nu^k_\deltat$ is the discretization of a time continuous \emph{Ornstein-Uhlenbeck}
process, defined as
\begin{equation}
	d\nu_t = - \nu_t \theta dt + \sigma \dbt.
	\label{eq:orn_uhl}
\end{equation}

@@

\subsection{Algorithms}
\label{subsec:algorithm}
To learn $V_{\theta}$ and $A_{\psi}$, variants of Q-learning for continuous~\cite{ddpg} and
discrete~\cite{dqn} action spaces are used. In details, $V_{\theta}$ and $A_{\psi}$ are learnt
using approximate version of the following bellman equation, which holds for the true $A^\pi_\deltat$
and $V^\pi_\deltat$,
\begin{align}
	V^{\pi}_\deltat(s) + \deltat A^{\pi}_\deltat(s, a) &= r\deltat + \gamma^{\deltat}  \E_{s'} V^{\pi}_\deltat(s')\label{eq:bellman_A}\\
	A^{\pi}_\deltat(s, \pi(s)) &= 0\label{eq:max_A}
\end{align}
for a greedy exploitation policy $\pi(s) = \text{argmax}_{a'}A^\pi_\deltat(s,
a')$, exactly implemented in the case of discrete actions, and approximated by
a neural network $\pi_\phi(s)$, trained to maximize (an approximation of)
$A^\pi_\deltat(s, \pi_\phi(s))$ in the continuous case.  \LB{Maybe say here that it's the same than DDPG?}Eq.~\eqref{eq:max_A}
is directly verified by $A_{\psi}$, owing to the reparametrization
$A_\psi(s, a) = \bar{A}_\psi(s, a) - \bar{A}_\psi(s, \pi(s))$, described
in~\ref{subsec:reparam}.  To approximately verify~\eqref{eq:bellman_A}, the
corresponding squared residual is minimized by an approximate gradient descent.
The corresponding update equations when learning from a transition $(s, a, r, s')$ of
an exploratory trajectory are
\begin{align}
	\delta Q_\deltat &\leftarrow A(s, a) \deltat - \left(r \deltat + \gamma^{\deltat} V(s') - V(s)\right)\\
	\theta_\deltat &\leftarrow \theta_\deltat + \eta^1_\deltat \partial_{\theta} V(s) \frac{\delta Q_\deltat}{\deltat}\\
	\psi_\deltat &\leftarrow \psi_\deltat + \eta^2_\deltat \partial_{\psi} A(s, a) \frac{\delta Q_\deltat}{\deltat}.
	\label{eq:approx_bellman_A}
\end{align}
\LB{I don't like the notations $\alpha^1$ and $\alpha^2$. What about $\alpha^A$ and $\alpha^V$? Easier to read} 
where $\eta$'s are learning rates. As we will see in the following section,
to obtain a well defined continuous time limit, the learning rates can be rewritten as
$\eta^i_\deltat = \alpha^i \deltat$, where $\alpha^1$ and $\alpha^2$ are independant
of $\deltat$. In practice, $A$ and $V$ are learned using
transitions sampled from a replay buffer instead of trajectory samples, to
prevent excessive correlation between training samples~\cite{dqn}.
\begin{algorithm}[ht]
	\caption{Deep Advantage Updating (Discrete actions)}
	\input{dauc.tex}
	\label{alg:dau}
\end{algorithm}
